---
title: "RWorksheet_Aguirre#4A"
author: "Ryza Faith"
date: "2024-10-14"
output: pdf_document
---
1A.
Describe the Data.

Shoe size is measured as a number, and most people in the data have shoe sizes between 8.5 and 10.5.  
Height shows more widely, but most people are between 64 and 72 inches tall.
```{r}
data <- data.frame(
  Shoe_Size = c(6.5, 9.0, 8.5, 8.5, 10.5, 7.0, 9.5, 9.0, 13.0, 7.5, 10.5, 8.5, 12.0, 10.5, 13.0, 11.5, 8.5, 5.0, 10.0, 6.5, 7.5, 8.5, 10.5, 8.5, 10.5, 11.0, 9.0, 13.0),
Height = c(66.0, 68.0, 64.5, 65.0, 70.0, 64.0, 70.0, 71.0, 72.0, 64.0, 74.75, 67.0, 71.0, 71.0, 77.0, 72.0, 59.0, 62.0, 72.0, 66.0, 64.0, 67.0, 73.0, 69.0, 72.0, 70.0, 69.0, 70.0),
Gender = c("F", "F", "F", "F", "M", "F", "F", "F", "M", "F", "M", "F", "M", "M", "M", "M", "F", "F", "M", "F", "F", "M", "M", "F", "M", "M", "M", "M")
)
data
```
1B.
```{r}
females <- subset(data, Gender == "F", select = c(Shoe_Size, Height))
females
```
1B.
```{r}
males <- subset(data, Gender == "M", select = c(Shoe_Size, Height))
males
```
1C.
```{r}
mean_shoe_size <- mean(data$Shoe_Size)
mean_shoe_size
```
1C.
```{r}
mean_height <- mean(data$Height)
mean_height
```

D.Is there a relationship between shoe size and height? Why?
- No, because the correlation coefficient between shoe size and height is close to 0, this would indicate no significant linear relationship between the two variables.

2. FACTOR

```{r}
Months <- c("March", "April", "JAnuary", "November", "January", "September", "October", "September", "November", "August", "January", "November", "November", "Febraury", "May", "Augsut")
factor_Months <- factor(Months)
factor_Months
```
3. 
```{r}
summary(Months)
summary(factor_Months)
```
4.Vector
```{r}
Directions <- c( "East", "West", "North")
Frequency <- c(1, 4, 3)
Directions
Frequency
```
4.Factor
```{r}
factor_data <- factor(Directions, levels = c("East", "West", "North"))
new_order_data <- factor(factor_data, levels = c("East", "West", "North"))

new_order_data

```
5.
```{r}
data <- read.table("~/DataScience/CS101/worksheet 4a/import_march.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE)
head(data)

```
6.
```{r}
exhaustive_search <- function(selected_number) 
  if (selected_number < 1 || selected_number > 50) {
    return("The number selected is beyond the range of 1 to 50")
  } else if (selected_number == 20) {
    return("TRUE")
  } else {
    return(as.character(selected_number))
  }

random_number <- sample(1:50, 1)
cat("The chosen number is:", random_number, "\n")
result <- exhaustive_search(random_number)
cat("Result:", result, "\n")

```
7.
```{r}
min_bills <- function(price) {
  bills <- c(1000, 500, 200, 100, 50)
  bill_count <- 0
  
  if (price %% 50 != 0) {
    return("Price must be a multiple of 50.")
  }
  
  for (bill in bills) {
    while (price >= bill) {
      price <- price - bill
      bill_count <- bill_count + 1
    }
  }
  
  return(bill_count)
}

price_of_snack <- 2700  

cat("Minimum number of bills needed:", min_bills(price_of_snack), "\n")

```
8A.
```{r}
data <- data.frame(
  Name = c("Annie", " Thea", "Steve", "Hanna"),
  Grade1 = c(85, 75, 75, 95),
  Grade2 = c(65, 75, 55, 75),
  Grade3 = c(85, 90, 80, 100),
  Grade4 = c(100, 90, 85, 90))
data
```
8B.
```{r}
for (i in 1:nrow(data)) {
  avg_score <- sum(data[i, 2:5]) / 4  
  
  if (avg_score >= 88.75) {
    cat(data$Name[i], "'s average grade this semester is", avg_score, "\n")
  }
}

```
8C.
```{r}
for (i in 1:nrow(data)) {
  avg_score <- sum(data[i, 2:5]) / 4  
  if (avg_score >= 88.75) {
    cat(data$Name[i], "'s average grade this semester is", avg_score, "\n")
  }
}

```
8D.
```{r}
results <- c()

for (i in 1:nrow(data)) {
  highest_score <- 0  
  
  for (j in 2:ncol(data)) {
    if (data[i, j] > highest_score) {
      highest_score <- data[i, j]  
    }
  }
  
  if (highest_score > 90) {
    results <- c(results, paste(data$Name[i], "’s highest grade this semester is", highest_score))
  }
}

cat(results, "\n")

```

